README

### 2022-09-28 rmd/supporting_info.Rmd
edit: misc

---
output:
  bookdown::pdf_document2:
    fig_caption: true
    number_sections: false
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
mainfont: "Times New Roman"
fontsize: 11pt
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{amsmath,fancyhdr, caption}
  - \captionsetup{width=1.3\linewidth, labelfont={bf}}
  - \pagestyle{fancy}
  - \fancyhead[L]{}
  - \fancyhead[C]{}
  - \setlength{\headheight}{13.7pt}
bibliography: reference.bib
csl: https://www.zotero.org/styles/pnas
---

@@

```{=tex}
\newpage
\pagenumbering{arabic}
\def\figurename{Figure S\hspace{-.4ex}}
\def\tablename{Table S\hspace{-.4ex}}
```
# Methods

@@

First, we analyzed a two-species community to gain insights into the model behavior and to guide the N-species analysis. Specifically, we examined the release effect with $20$ parameter values of intrinsic population growth rate $r_i$ ($0.5$ to $3.5$ with an equal interval) and carrying capacity $K$ ($50$ to $500$ with an equal interval), resulting in $400$ parameter combinations. In a single-species Ricker model, the chosen range of $r_i$ can generate stable equilibrium, damped oscillations, limit cycle, and chaos (XXX). For simplicity, we assumed $r_1 = r_2$. We crossed this parameter setup with environmental stochasticity ($\sigma_{\varepsilon} =$ `r unique(list_param[[1]]$sd_env)`) and competition strength ($\alpha_{ij} =$ `r unique(list_param[[1]]$alpha)`) and generated `r nrow(list_param[[1]])` simulation scenarios ($400 \times 2 \times 2$).

@@

```{=tex}
\begin{align}
\begin{split}
R'_s &= \eta_{w,s'} R_w\\
\sum_{s'}^{S_w + 1} \eta_{w,s'} &= 1
(\#eq:effr)
\end{split}
\end{align}
```
In Equation \@ref(eq:effr), $\eta_{w,s'}$ represents the effective proportion of total release affecting the community at site $s'$ within watershed $w$. The summation of $\eta_{w,s'}$ over $S_w+1$ is assumed to be unity because a non-negligible portion of release may escape into unsurveyed areas. Since the values of $\eta_{w,s'}$ are not available in our dataset, we estimated them through a stochastic search while fitting the regression model to the data within a Bayesian framework (see **model fitting**). Other site-level predictors include upstream watershed area (log-transformed), air temperature, precipitation, and forest land use. Urban and agricultural land use was omitted because of either a limited value range (**Figure S**\@ref(fig:fig-env)) or a strong correlation with forest land use (**Figure S**\@ref(fig:fig-cor)). The watershed-specific intercept was related to watershed-level predictors as:

```{=tex}
\begin{align}
\begin{split}
\gamma_{0,w} &\sim \text{Normal}(\theta_{\gamma,w},\sigma_{\gamma}^2)\\
\theta_{\gamma,w} &= \gamma'_0 + \gamma'_1 x'_{w}
\end{split}
\end{align}
```
$\gamma'_{0}$ is the global intercept and $\gamma'_1$ is the regression coefficient of ocean productivity $\gamma'_{w}$ (chlorophyll *a* concentration; averaged for 2002 -- 2019). Ocean productivity was included because the majority of the observed species use marine habitats at a certain life stage (i.e., diadromous). The parameter $\sigma_{\gamma}$ accounts for random variation among watersheds that the watershed-level predictors cannot capture.

Since the CV is an additive function of mean $\mu$ and SD $\sigma$ in a log scale, we can derive the effects of environmental factors on the CV from Equation \@ref(eq: model):

@@

**Model fitting.** We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned vague or weakly informative priors to parameters (**Table S**\@ref(tab:tab-prior)). Four Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. The total MCMC iterations ranged from `r min(df_mcmc$n_total_mcmc)` to `r max(df_mcmc$n_total_mcmc)`, in which MCMC samples were saved every `r unique(df_mcmc$n_thin)` steps for the calculation of posterior distributions after the initial `r max(df_mcmc$n_burn)` burn-in period. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached $<1.1$ [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Tables S**\@ref(tab:tab-reg1) -- S\@ref(tab:tab-reg3).

@@

knitr::kable(df_prior,
             format = "markdown",
             caption = "Prior distributions used in the Bayesian models. See Durante [@duranteNoteMultiplicativeGamma2017] for multiplicative gamma priors. The parameter $\\pmb{\\phi}$ is a vector of prior scales ($\\pmb{\\phi}=2.5,...,2.5$).")

@@

## Theoretical prediction (2-species community)

```{r fig-2sp, fig.width=11, fig.height=8, fig.cap = paste0("Theoretical predictions in a 2-species community. **(A)** Spearman's rank correlations between intentional release and community dynamics in a strong competition scenario ($\\bar{\\alpha}=$ ", max(list_param[[1]]$alpha) ,"). Rows distinguish different summary statistics of community dynamics (top: CV, middle: mean, bottom: SD) while columns distinguish deterministic ($\\sigma_{\\varepsilon}=$ ", min(list_param[[1]]$sd_env) ,") and stochastic simulations ($\\sigma_{\\varepsilon}=$ ", max(list_param[[1]]$sd_env) ,"). Cells indicate 400 combinations of carrying capacity and intrinsic growth rate (20 values each). Colors are proportional to Spearman's rank correlations. Dots indicate parameter combinations that are used in the whole community simulation. **(B)** Examples of community dynamics with and without intentional release. Community dynamics diverge at the broken vertical lines, the timestep at which intentional release begins. Rows distinguish intrinsic growth rates, while columns distinguish deterministic and stochastic simulations. Colors indicate release levels, and line types differ in carrying capacities.")}

@@

  cap[i] <- paste0("Theoretical predictions for the release effect on community dynamics. Rows represent different response variables, and columns show distinct simulation scenarios with different strengths of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Other parameters are: ",
                       "number of species $S$ = ", df_param$n_species[i], "; intrinsic growth rate of an enhanced species $r_1$ = ", df_param$r1[i], "; maximum intrinsic growth rate of unenhanced species $r_{max}$ = ", df_param$r_max[i], "; environmental stochasticity $\\sigma_{\\varepsilon}$ = ", df_param$sd_env[i], "; carrying capacity $K$ = ", df_param$k[i], ".")

  knit_chunk <- paste0("## Theoretical prediction (",
                       "$r_1$ = `r df_param$r1[",i,"]`, ",
                       "$K$ = `r df_param$k[",i,"]`) \n",
                       "\n```{r fig-sim", i,", fig.width = 10, fig.height = 7, fig.cap = cap[", i, "]}
                       \n\n
                       list_g_theory[[",i,"]]
                       \n\n
                       ```\n",
                        "\n\\pagebreak\n")

@@

```{r fig-int, fig.width=11, fig.height=11, fig.cap="Estimates of competition coefficients $\\alpha_{ij}$ in a multi-species Ricker model. Panels distinguish sites, and values in each cell indicate median estimates of $\\alpha_{ij}$ for a given pair of species. Cells are colored in proportion to $\\alpha_{ij}$."}

source(here::here("code/si_figure_species_int.R"))
print(g_alpha_site)
```

>>>>

---
output:
  bookdown::pdf_document2:
    fig_caption: true
    number_sections: false
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
mainfont: "Times New Roman"
fontsize: 11pt
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{amsmath,fancyhdr, caption}
  - \captionsetup{width=1.3\linewidth, labelfont={bf}}
bibliography: reference.bib
csl: https://www.zotero.org/styles/pnas
---

@@

```{=tex}
\newpage
\pagestyle{fancy}
\fancyhead[L]{Terui et al.}
\fancyhead[C]{}
\setlength{\headheight}{13.7pt}
\pagenumbering{arabic}
\def\figurename{Figure S\hspace{-.4ex}}
\def\tablename{Table S\hspace{-.4ex}}
```
# Methods

@@

First, we analyzed a two-species community to gain insights into the model behavior and to guide the N-species analysis. Specifically, we examined the release effect with $20$ parameter values of intrinsic population growth rate $r_i$ ($0.5$ to $3.5$ with an equal interval) and carrying capacity $K$ ($50$ to $500$ with an equal interval), resulting in $400$ parameter combinations. In a single-species Ricker model, the chosen range of $r_i$ can generate stable equilibrium, damped oscillations, limit cycle, and chaos [@sunDynamicBehaviorsRicker2000]. For simplicity, we assumed $r_1 = r_2$. We crossed this parameter setup with environmental stochasticity ($\sigma_{\varepsilon} =$ `r unique(list_param[[1]]$sd_env)`) and competition strength ($\alpha_{ij} =$ `r unique(list_param[[1]]$alpha)`) and generated `r nrow(list_param[[1]])` simulation scenarios ($400 \times 2 \times 2$).

@@

```{=tex}
\begin{align}
\begin{split}
R'_s &= \eta_{w,s'} R_w\\
\sum_{s'}^{S_w + 1} \eta_{w,s'} &= 1
(\#eq:effr)
\end{split}
\end{align}
```
In Equation \@ref(eq:effr), $\eta_{w,s'}$ represents the effective proportion of total release affecting the community at site $s'$ within watershed $w$. The summation of $\eta_{w,s'}$ over $S_w+1$ is assumed to be unity because a non-negligible portion of release may escape into unsurveyed areas. Since the values of $\eta_{w,s'}$ are not available in our dataset, we estimated them through a stochastic search while fitting the regression model to the data within a Bayesian framework (see **model fitting**). Other site-level predictors include upstream watershed area (log-transformed), air temperature, precipitation, and forest land use. Urban and agricultural land use was omitted because of either a limited value range (**Figure S**\@ref(fig:fig-env)) or a strong correlation with forest land use (**Figure S**\@ref(fig:fig-cor)). The watershed-specific intercept was related to a watershed-level predictor as:

```{=tex}
\begin{align}
\begin{split}
\gamma_{0,w} &\sim \text{Normal}(\theta_{\gamma,w},\sigma_{\gamma}^2)\\
\theta_{\gamma,w} &= \gamma'_0 + \gamma'_1 x'_{w}
\end{split}
\end{align}
```
$\gamma'_{0}$ is the global intercept and $\gamma'_1$ is the regression coefficient of ocean productivity $\gamma'_{w}$ (chlorophyll *a* concentration; averaged for 2002 -- 2019). Ocean productivity was included because the majority of the observed species use marine habitats at a certain life stage (i.e., diadromous). The parameter $\sigma_{\gamma}$ accounts for random variation among watersheds that the watershed-level predictor cannot capture.

Since the CV is an additive function of mean $\mu$ and SD $\sigma$ in a log scale, we can derive the effects of environmental factors on the CV from Equation \@ref(eq:model):

@@

**Model fitting.** We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned vague or weakly informative priors to parameters (**Table S**\@ref(tab:tab-prior)). Four Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. The total MCMC iterations ranged from `r min(df_mcmc$n_total_mcmc)` to `r max(df_mcmc$n_total_mcmc)`, in which MCMC samples were saved every `r unique(df_mcmc$n_thin)` steps for the calculation of posterior distributions after the initial `r max(df_mcmc$n_burn)` burn-in period. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached $<1.1$ [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Tables S**\@ref(tab:tab-ssm-ar) -- **S**\@ref(tab:tab-reg3).

@@

knitr::kable(df_prior,
             format = "markdown",
             caption = "Prior distributions used in the Bayesian models. See Durante [@duranteNoteMultiplicativeGamma2017] for multiplicative gamma priors. The parameter $\\pmb{\\omicron}$ is a vector of prior scales ($\\pmb{\\omicron}=2.5,...,2.5$).")

@@

## Theoretical prediction (two-species community)

```{r fig-2sp, fig.width=12, fig.height=8, fig.cap = paste0("Theoretical predictions in a two-species community. **(A)** Spearman's rank correlations between intentional release and community dynamics in a strong competition scenario ($\\bar{\\alpha}=$ ", max(list_param[[1]]$alpha) ,"). Rows distinguish different summary statistics of community dynamics (top: CV, middle: mean, bottom: SD) while columns distinguish deterministic ($\\sigma_{\\varepsilon}=$ ", min(list_param[[1]]$sd_env) ,") and stochastic simulations ($\\sigma_{\\varepsilon}=$ ", max(list_param[[1]]$sd_env) ,"). Cells indicate 400 combinations of carrying capacity and intrinsic growth rate (20 values each). Colors are proportional to Spearman's rank correlations. Dots indicate parameter combinations that are used in the whole community simulation. **(B)** Examples of community dynamics with and without intentional release. Community dynamics diverge at the broken vertical lines, the timestep at which intentional release begins. Rows distinguish intrinsic growth rates, while columns distinguish deterministic and stochastic simulations. Colors indicate release levels, and line types differ in carrying capacities.")}

@@

  cap[i] <- paste0("Theoretical predictions for the release effect in a ten-species community. Rows represent different response variables, and columns show distinct simulation scenarios with different strengths of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Other parameters are: ",
                       "number of species $S$ = ", df_param$n_species[i], "; intrinsic growth rate of an enhanced species $r_1$ = ", df_param$r1[i], "; maximum intrinsic growth rate of unenhanced species $r_{max}$ = ", df_param$r_max[i], "; environmental stochasticity $\\sigma_{\\varepsilon}$ = ", df_param$sd_env[i], "; carrying capacity $K$ = ", df_param$k[i], ".")

  knit_chunk <- paste0("## Theoretical prediction (",
                       "$r_1$ = `r df_param$r1[",i,"]`, ",
                       "$K$ = `r df_param$k[",i,"]`) \n",
                       "\n```{r fig-sim", i,", fig.width = 10, fig.height = 8.5, fig.cap = cap[", i, "]}
                       \n\n
                       list_g_theory[[",i,"]]
                       \n\n
                       ```\n",
                        "\n\\pagebreak\n")

@@

```{r fig-int, fig.width=7, fig.height=7, fig.cap="Estimates of competition coefficients $\\alpha_{ij}$ in a multi-species Ricker model. Panels distinguish sites, and values in each cell indicate median estimates of $\\alpha_{ij}$ for a given pair of species. Cells are colored in proportion to $\\alpha_{ij}$."}

source(here::here("code/si_figure_species_int.R"))
knitr::include_graphics(here::here("output/figure_alpha_site.pdf"))
```
